Sparse Function-space Representation \\ of Neural Networks

### 2023-05-17 paper/main.tex
updates in sec. 1.1

%Current state of affairs 
Recent techniques (\eg, \cite{ritter2018kfac,khan2019approximate,daxberger2021laplace,fortuin2021bayesian,immer2021scalable}) apply the Laplace-GGN approximation\todo{Is Laplace-GGN a well-known acronym?} to convert a trained neural networks into a Bayesian neural network, that can provide uncertainty without sacrificing additional resources to training \cite{foong2019between}. Furthermore, the resultant weight-space posterior can be converted to the function-space as shown in \cite{khan2019approximate, immer2021improving}. The function-space representation allows for a principled mathematical approach for analyzing the behaviour \cite{cho2009kernel,meronen2020stationary}, performing probabilistic inference \cite{khan2019approximate}, and quantifying uncertainty in neural networks \cite{foong2019between}. These methods rely on the linearization of the neural network and the resultant neural tangent kernel (NTK, \cite{jacot2018neural}). The neural network is characterized in function-space by its first two moment functions, a mean function and covariance function (or kernel)---defining a Gaussian process (GP, \cite{rasmussen2006gaussian}). GPs provide a widely-used probabilistic toolkit with principled uncertainty estimates. They serve as a standard surrogate model for Bayesian optimization \citep{garnett_bayesoptbook_2022} and are effective in model-based reinforcement learning \citep{deisenroth2011pilco} with theoretical guarantees on regret bounds \citep{srinivas2009gaussian}.  
%Yet many problems lie in high dimensional input space; for example, images are where GPs cannot learn representations. In such scenarios such as in many reinforcement learning environments neural networks are used as the surrogate model. However, uncertainty is still essential to ensure effective exploration for sequential algorithms. Successful approaches have attempted to blend neural networks with uncertainty estimates around predictions, allowing for sophisticated exploration strategies. However, there has been limited use of hybrid models that possess the feature representation ability of neural networks but also attractive the properties of GPs, such a hybrid method we propose in this paper.

@@

% Introduce the laplace GGN + shortcomings
\textbf{Function-space methods}
Function-space perspectives on uncertainty in neural networks often use a Laplace-GGN approximation \citep{daxberger2021laplace}, linearizing a trained neural network around optimal weights. \todo{Could modify the sentence a bit, maybe "The Laplace-GGN approximates the Hessian by using..."}The Hessian can be efficiently approximated using the generalized Gauss--Newton approximation (GGN, \cite{botev2017practical}). While efficient, it suffers from cubic scaling in parameter count, necessitating approximations like Kronecker factorisation \cite{martens2015optimizing, ritter2018kfac}. \todo{Reads a bit repetitive to start two sentences in a row by This, could maybe be combined in one.} This linear model (linear in the weights) refines predictions and provides uncertainty estimates \citep{immer2021scalable}. This model has a convenient interpretation as a GP \cite{immer2021scalable, khan2019approximate, maddox2021fast}.
%
However, the GP's cubic $\mathcal{O}(n^3)$ scaling with data points $n$ requires costly approximations, often resorting to using data subsets \cite{immer2021scalable}. In the GP community, sparse approximations have mitigated this scaling issue (\eg, \cite{titsias2009variational,hensman2013gaussian}), but combining neural network linearization with sparse methods remains \todo{Maybe unexplored instead of unclear?} unclear without resorting to subset methods \todo{Could add something on subset methods here, for example 'which do not retain information from the full dataset' or something like that}. Our work addresses this by utilizing the GP's dual parameters \cite{csato2002sparse}, previously applied to non-conjugate likelihood models \cite{adam2021dual}.

>>>>

%Current state of affairs 
Recent techniques (\eg, \cite{ritter2018kfac,khan2019approximate,daxberger2021laplace,fortuin2021bayesian,immer2021scalable}) apply a Laplace-GGN approximation\todo{Is Laplace-GGN a well-known acronym?} to convert trained neural networks into Bayesian neural networks, that can provide uncertainty without sacrificing additional resources to training \cite{foong2019between}. Furthermore, the resultant weight-space posterior can be converted to the function-space as shown in \cite{khan2019approximate, immer2021improving}. The function-space representation allows for a principled mathematical approach for analyzing the behaviour \cite{cho2009kernel,meronen2020stationary}, performing probabilistic inference \cite{khan2019approximate}, and quantifying uncertainty in neural networks \cite{foong2019between}. These methods rely on the linearization of the neural network and the resultant neural tangent kernel (NTK, \cite{jacot2018neural}). The neural network is characterized in function-space by its first two moment functions, a mean function and covariance function (or kernel)---defining a Gaussian process (GP, \cite{rasmussen2006gaussian}). GPs provide a widely-used probabilistic toolkit with principled uncertainty estimates. They serve as a standard surrogate model for Bayesian optimization \citep{garnett_bayesoptbook_2022} and are effective in model-based reinforcement learning \citep{deisenroth2011pilco} with theoretical guarantees on regret bounds \citep{srinivas2009gaussian}.  
%Yet many problems lie in high dimensional input space; for example, images are where GPs cannot learn representations. In such scenarios such as in many reinforcement learning environments neural networks are used as the surrogate model. However, uncertainty is still essential to ensure effective exploration for sequential algorithms. Successful approaches have attempted to blend neural networks with uncertainty estimates around predictions, allowing for sophisticated exploration strategies. However, there has been limited use of hybrid models that possess the feature representation ability of neural networks but also attractive the properties of GPs, such a hybrid method we propose in this paper.

@@

% Introduce the laplace GGN + shortcomings
\textbf{Function-space methods}
Function-space perspectives on uncertainty in neural networks often use a Laplace-GGN approximation \citep{daxberger2021laplace}, that linearize a trained neural network around MAP weights and approximates the neural network's Hessian using the generalized Gauss--Newton approximation (GGN, \cite{botev2017practical}). 
%\todo{Could modify the sentence a bit, maybe "The Laplace-GGN approximates the Hessian by using..."}
%The Hessian can be efficiently approximated using the generalized Gauss--Newton approximation . 
While efficient, it suffers from cubic scaling in parameter count, necessitating approximations like Kronecker factorisation \cite{martens2015optimizing, ritter2018kfac}. \todo{Reads a bit repetitive to start two sentences in a row by This, could maybe be combined in one.} This linear model (linear in the weights) refines predictions and provides uncertainty estimates \citep{immer2021scalable}. This model has a convenient interpretation as a GP \cite{immer2021scalable, khan2019approximate, maddox2021fast}.
%
However, the GP's cubic $\mathcal{O}(n^3)$ scaling with data points $n$ requires costly approximations, often resorting to using data subsets \cite{immer2021scalable}. In the GP community, sparse approximations have mitigated this scaling issue (\eg, \cite{titsias2009variational,hensman2013gaussian}), but combining neural network linearization with sparse methods remains \todo{Maybe unexplored instead of unclear?} unclear without resorting to subset methods \todo{Could add something on subset methods here, for example 'which do not retain information from the full dataset' or something like that}. Our work addresses this by utilizing the GP's dual parameters \cite{csato2002sparse}, previously applied to non-conjugate likelihood models \cite{adam2021dual}.
